---
title: "NIH Human Monocyte RNA-Seq analysis"
author: "Laura Blanton"
output: html_document
---

####Gene annotations
```{r}
geneAnno <- read.delim(file="/sca-immune/Annotations/geneAnno_proteinCoding_lncRNA_v107_20221021.txt", stringsAsFactors = FALSE)[,c(2,1,3:6,8,14,9)]
colnames(geneAnno)[c(1,4,5,6)] <- c("Gene","chr","start","stop")
geneAnno <- geneAnno[! duplicated(geneAnno$gene_name.107) & ! geneAnno$gene_name.107 %in% c("") & ! duplicated(geneAnno$Gene),]
geneAnno_x <- geneAnno[geneAnno$chr == "chrX",]
geneAnno_y <- geneAnno[geneAnno$chr == "chrY",]
X_genes_all <-as.character(geneAnno[grep("chrX", geneAnno$chr), ]$Gene)
Y_genes_all <-as.character(geneAnno[grep("chrY", geneAnno$chr), ]$Gene)
sexChrom_genes <- c(X_genes_all,Y_genes_all)
PAR1_genes <- geneAnno_x[geneAnno_x$start < 2691188,]$Gene
PAR2_genes <- c("SPRY3","VAMP7","IL9R","WASH6P")
PAR_genes_all <- c(PAR1_genes,PAR2_genes)
NPX_genes <- X_genes_all[! X_genes_all %in% PAR_genes_all]
NPY_genes <- Y_genes_all[! Y_genes_all %in% PAR_genes_all]
autosome_genes <- geneAnno[! geneAnno$chr %in% c("chrX","chrY"),"Gene"]
X_ancestral_genes = c("SOX3", "ZFX", "DDX3X", "KDM6A","KDM5C","USP9X","EIF1AX","RPS4X","AMELX","TXLNG","TMSB4X","NLGN4X","TBL1X","PRKX")
Y_ancestral_genes = c("SRY", "ZFY", "DDX3Y", "UTY","KDM5D","USP9Y","EIF1AY","RPS4Y1","RPS4Y2", "AMELY","TXLNGY","TMSB4Y","NLGN4Y","TBL1Y","PRKY")
XY_ancestral_genes = c("SOX3", "ZFX", "DDX3X", "KDM6A","KDM5C","USP9X","EIF1AX","RPS4X","AMELX","TXLNG","TMSB4X","NLGN4X","TBL1X","PRKX","SRY", "ZFY", "DDX3Y", "UTY","KDM5D","USP9Y","EIF1AY","RPS4Y1","RPS4Y2", "AMELY","TXLNGY","TMSB4Y","NLGN4Y","TBL1Y","PRKY")

```

####Load metadata 
```{r}
library("tximport")
library("readr")

setwd("") #set path to working directory
dir <-  #set path to directory with metadata
myPath <-  #set path to directory with kallisto results
  
#Set up annotation
annofile <-read.delim(file="annotation_with_ERCC.txt" ,sep = " ") #path to annotation_with_ERCC.txt
tx2gene <- data.frame("TXNAME"=annofile$transcript_id, "GENEID"=annofile$gene_name)

#Set up file lists and metadata
samples <- read.delim(file.path(dir, "monocyte_metadata.csv"), sep = "\t", header = TRUE)
monofiles <- file.path(dir, "monocyte_results", samples$Kallisto_directory, "abundance.tsv")
names(monofiles) <- samples$Kallisto_directory

#Set up monocyte sample metadata
colData <- rread.delim(file.path(dir, "monocyte_metadata.csv"), sep = "\t", header = TRUE, row.names="SampleID")
colData <- colData[,c("Sex","Karyotype", "X_count", "Y_count", "Sex_chrom_count", "GTC_Run_ID")]
```

####Import monocyte RNA-Seq data
```{r}
#####This is code for importing kallisto results generated from raw fastq files; the raw fastq files are accessible at dbGaP accession # phs002481. From the raw data, run the following kallisto command: 
# kallisto quant -i KALLISTO_INDEX_FILE -t 16 --bias --plaintext -o  /kallisto_OUTPUT_FOLDER/sampleName/ file1fastq.gz file2fastq.gz 

# all monocyte files
all(file.exists(monofiles))
mono.txi <- tximport(monofiles, type = "kallisto", tx2gene = tx2gene)
#save(mono.txi, file="mono.txi.rda")

# 2X+ karyotypes
samples_2Xup <- subset(samples, samples$X_count >= 2)
monofiles_2Xup <- file.path(dir, "monocyte_results", samples_2Xup$Kallisto_directory, "abundance.tsv")
names(monofiles_2Xup) <- samples_2Xup$Kallisto_directory
all(file.exists(monofiles_2Xup))
mono.2Xup.txi <- tximport(monofiles_2Xup, type = "kallisto", tx2gene = tx2gene)
#save(mono.2Xup.txi, file="mono.2Xup.txi.rda")

# Y karyotypes
samples_Y <- subset(samples, samples$Y_count >= 1)
monofiles_Y <- file.path(dir, "monocyte_results", samples_Y$Kallisto_directory, "abundance.tsv")
names(monofiles_Y) <- samples_Y$Kallisto_directory
all(file.exists(monofiles_Y))
mono.Y.txi <- tximport(monofiles_Y, type = "kallisto", tx2gene = tx2gene)
#save(mono.Y.txi, file="mono.Y.txi.rda")

# 0Y karyotypes
samples_y.absent <- subset(samples, samples$Y_count == 0)
monofiles_y.absent <- file.path(dir, "monocyte_results", samples_y.absent$Kallisto_directory, "abundance.tsv")
names(monofiles_y.absent) <- samples_y.absent$Kallisto_directory
all(file.exists(monofiles_y.absent))
mono.y.absent.txi <- tximport(monofiles_y.absent, type = "kallisto", tx2gene = tx2gene)
#save(mono.y.absent.txi, file="mono.y.absent.txi.rda")

# XY and XXY only
samples_1Y <- subset(samples, samples$Y_count == 1)
monofiles_1Y <- file.path(dir, "monocyte_results", samples_1Y$Kallisto_directory, "abundance.tsv")
names(monofiles_1Y) <- samples_1Y$Kallisto_directory
all(file.exists(monofiles_1Y))
mono.1Y.txi <- tximport(monofiles_1Y, type = "kallisto", tx2gene = tx2gene)
#save(mono.1Y.txi, file="mono.n1Y.txi.rda")

# 1X karyotypes with at least 1Y (XY and XYY only)
samples_1X <- subset(samples, samples$X_count == 1 & samples$Y_count >= 1)
monofiles_1X <- file.path(dir, "monocyte_results", samples_1X$Kallisto_directory, "abundance.tsv")
names(monofiles_1X) <- samples_1X$Kallisto_directory
all(file.exists(monofiles_1X))
mono.1X.txi <- tximport(monofiles_1X, type = "kallisto", tx2gene = tx2gene)
#save(mono.1X.txi, file="mono.1X.txi.rda")
```

####Find expressed genes - code adapted from Adrianna San Roman
```{r}
####Find expressed genes
txi.TPM <- mono.txi$abundance

#Get list of expressed genes in the cell type of interest:
XX_sample_names <- colData[colData$Karyotype == "46XX", "Kallisto_directory"]
XY_sample_names <- colData[colData$Karyotype == "46XY", "Kallisto_directory"]
tpm1_genes_xx_xy_median <- function(tximport_file, XX_samples, XY_samples) {
  kallisto_output_TPM_XY <- tximport_file$abundance[,c(XY_samples)]
  kallisto_output_TPM_XX <- tximport_file$abundance[,c(XX_samples)]
  median_TPM_XY <- apply(kallisto_output_TPM_XY,1,median)
  median_TPM_XX <- apply(kallisto_output_TPM_XX,1,median)
  #Put all of the medians together in a dataframe of expression per karyotype - numeric
  expression_perKaryotype <- data.frame(median_TPM_XY, median_TPM_XX)
  #Isolate the rows of the expression_TPM dataset that have at least one karyotype passing the threshold
  expressedGenes <- expression_perKaryotype[expression_perKaryotype$median_TPM_XY >= 1 | expression_perKaryotype$median_TPM_XX >= 1,]
  #Restrict by gene annotation file
  expressedGenes <- expressedGenes[rownames(expressedGenes) %in% geneAnno$Gene,]
  write.table(expressedGenes, file=paste0("expressedGenes_TPM_",tximport_file,".txt"),quote=FALSE,sep="\t")
  #See how many genes are expressed in at least one karyotype
  num_expressedGenes <- dim(expressedGenes)[1]
  print(paste("Number of 'expressed' genes with median TPM >= 1 in either XX or XY samples: ", as.character(num_expressedGenes), sep=""))

  #Find expressed sex chromosome genes
  expSexChromGenes <- expressedGenes[rownames(expressedGenes) %in% sexChrom_genes,]
  print(paste("Number of 'expressed' sex chromosome genes: ", as.character(dim(expSexChromGenes)[1]), sep=""))
  #write.table(expSexChromGenes, file="expressedSexChromGenes_TPM.txt",quote=FALSE,sep="\t")

  #Find expressed Y chromosome genes
  expYGenes <- expressedGenes[rownames(expressedGenes) %in% Y_genes_all,]
  print(paste("Number of 'expressed' Y chromosome genes: ", as.character(dim(expYGenes)[1]), sep=""))
  #write.table(expYGenes, file="expressedYgenes_TPM.txt",quote=FALSE,sep="\t")

  #Find expressed X chromosome genes
  expXGenes <- expressedGenes[rownames(expressedGenes) %in% X_genes_all,]
  print(paste("Number of 'expressed' X chromosome genes: ", as.character(dim(expXGenes)[1]), sep=""))
  #write.table(expXGenes, file="expressedXgenes_TPM.txt",quote=FALSE,sep="\t")

  #Find expressed PAR genes
  expPARGenes <- expressedGenes[rownames(expressedGenes) %in% PAR_genes_all,]
  print(paste("Number of 'expressed' PAR genes (small set): ", as.character(dim(expPARGenes)[1]), sep=""))
  #write.table(expPARGenes, file="expressedPARgenes_TPM.txt",quote=FALSE,sep="\t")

  #Find expressed autosome genes
  expAutoGenes <- expressedGenes[rownames(expressedGenes) %in% autosome_genes,]
  print(paste("Number of 'expressed' autosome genes: ", as.character(dim(expAutoGenes)[1]), sep=""))
  #write.table(expAutoGenes, file="expressedAutosomeGenes_TPM.txt",quote=FALSE,sep="\t")

  myList <- list("expressedGenes" = rownames(expressedGenes), "expSexChromGenes" = rownames(expSexChromGenes),
                 "expYGenes" = rownames(expYGenes), "expXGenes" = rownames(expXGenes), "expPARGenes" = rownames(expPARGenes),
                 "expAutoGenes" = rownames(expAutoGenes))

  return(myList)
}
 
mono_expressed <- tpm1_genes_xx_xy_median(tximport_file = mono.txi,XX_samples = XX_sample_names, XY_samples = XY_sample_names )
save(cd4_expressed, file="mono_expressed.rda")

expressedGenes <- mono_expressed$expressedGenes
```


####Make DESeq2 object: X or Y dosage model
```{r}
####Make DESeq2 object from imported kallisto files
library(DESeq2)
colnames(mono.txi$abundance) <- NULL
colnames(mono.txi$counts) <- NULL
colnames(mono.txi$length) <- NULL

mono.dds <- DESeqDataSetFromTximport(mono.txi, colData=colData, design=~ GTC_Run_ID + Y_count + X_count)
mono.dds <- DESeq(mono.dds)
save(mono.dds, file="mono.dds.rda")
```

####Make model with only Y karyotypes
```{r}
load(file="mono.Y.txi.rda")
colData.y <- subset(colData, colData$Y_count >=1)

colnames(mono.Y.txi$abundance) <- NULL
colnames(mono.Y.txi$counts) <- NULL
colnames(mono.Y.txi$length) <- NULL

mono.y.dds <- DESeqDataSetFromTximport(mono.Y.txi, colData=colData.y, design=~ GTC_Run_ID + Y_count + X_count)
mono.y.dds <- DESeq(mono.y.dds)
save(mono.y.dds, file="mono.y.dds.rda")
```

####Make model with only 0Y karyotypes
```{r}
load(file="mono.noY.txi.rda")
colData.noY <- subset(colData, colData$Y_count ==0)

colnames(mono.y.absent.txi$abundance) <- NULL
colnames(mono.y.absent.txi$counts) <- NULL
colnames(mono.y.absent.txi$length) <- NULL

mono.y.absent.dds <- DESeqDataSetFromTximport(mono.y.absent.txi, colData=colData.y.absent, design=~ GTC_Run_ID + X_count)
mono.y.absent.dds <- DESeq(mono.y.absent.dds)
save(mono.y.absent.dds, file="mono.y.absent.dds.rda")
```

####Make model with only 1Y karyotypes
```{r}
load(file="mono.n1Y.txi.rda")
colData.1Y <- subset(colData, colData$Y_count ==1)

colnames(mono.1Y.txi$abundance) <- NULL
colnames(mono.1Y.txi$counts) <- NULL
colnames(mono.1Y.txi$length) <- NULL

mono.1y.dds <- DESeqDataSetFromTximport(mono.1Y.txi, colData=colData.1Y, design=~ GTC_Run_ID + X_count)
mono.1y.dds <- DESeq(mono.1y.dds)

save(mono.1y.dds, file="mono.1y.dds.rda")
```

####Make model with 1X and >=1Y karyotypes
```{r}
load(file="mono.1X.txi.rda")
colData.1X <- subset(colData, colData$X_count ==1 & colData$Y_count >= 1)

colnames(mono.1X.txi$abundance) <- NULL
colnames(mono.1X.txi$counts) <- NULL
colnames(mono.1X.txi$length) <- NULL

mono.1x.dds <- DESeqDataSetFromTximport(mono.1X.txi, colData=colData.1X, design=~ GTC_Run_ID + Y_count)
mono.1x.dds <- DESeq(mono.1x.dds)

save(mono.1x.dds, file="mono.1x.dds.rda")
```

####Load model - if already created
```{r}
###After DESeq2 models have been generated and saved, the following code can be used to introduce the models into a new R environment 

source("tximport_supplement_fixed.R")
library("tximport")
library("readr")

setwd("") <- #set path to working directory

#Set up monocyte sample metadata
colData <- read.csv("monocyte_metadata.csv",sep="\t", row.names="SampleID")
colData <- colData[,c("Sex","Karyotype", "X_count", "Y_count", "Sex_chrom_count", "GTC_Run_ID")]

#Read in monocyte sample metadata
load(file="mono.dds.rda")
load(file="mono.txi.rda")

load(file="mono.y.dds.rda")
load(file="mono.y.absent.dds.rda")
load(file="mono.1y.dds.rda")
load(file="mono.1x.dds.rda")

load(file="mono_expressed.rda")
expressedGenes <- mono_expressed$expressedGenes
```

####Find significant DE genes
```{r}
source("/sca-immune/Rfunctions/Function_GenomeWideResponse.R")
source("/sca-immune/Rfunctions/Function_SignificantResponse.R")
source("/sca-immune/Rfunctions/Function_AutosomalResponse_221108.R")

####Monocyte: Look at significantly X-responsive genes####
#####Responsive to X dosage
x.response <- GenomeWideResponse(mono.dds, "X_count", expressedGenes) 
x.response.sig <- SignificantResponse(mono.dds, "X_count", expressedGenes)
x.response.auto <- AutosomalResponse(x.response)
x.response.sig.auto <- AutosomalResponse(x.response.sig)

#write.table(x.response.auto, "monocyte_autosomal_Xresponse.txt", quote = F, sep = "\t")
#write.table(x.response.sig.auto, "monocyte_autosomal_sig_Xresponse.txt", quote = F, sep = "\t")

Xresponsive_pos <- subset(x.response.sig.auto, x.response.sig.auto$log2FoldChange >= 0)
Xresponsive_neg <- subset(x.response.sig.auto, x.response.sig.auto$log2FoldChange <= 0)

####Responsive to Y dosage
y.response <- GenomeWideResponse(mono.dds, "Y_count", expressedGenes) 
y.response.sig <- SignificantResponse(mono.dds, "Y_count", expressedGenes)
y.response.auto <- AutosomalResponse(y.response)
y.response.sig.auto <- AutosomalResponse(y.response.sig)

Yresponsive_pos <- subset(y.response.sig.auto, y.response.sig.auto$log2FoldChange >= 0)
Yresponsive_neg <- subset(y.response.sig.auto, y.response.sig.auto$log2FoldChange <= 0)

#write.table(y.response.auto, "monocyte_autosomal_Yresponse.txt", quote = F, sep = "\t")
#write.table(y.response.sig.auto, "monocyte_autosomal_sig_Yresponse.txt", quote = F, sep = "\t")

####Y-present karyotypes model (XY, XXY, XYY)
y.present.x.response <- GenomeWideResponse(mono.y.dds, "X_count", expressedGenes)
y.present.x.response.auto <- AutosomalResponse(y.present.x.response)

y.present.y.response <- GenomeWideResponse(mono.y.dds, "Y_count", expressedGenes)
y.present.y.response.auto <- AutosomalResponse(y.present.y.response)

#write.table(y.only.x.response.auto, "monocyte_autosomal_Ypresent_Xresponse.txt", quote = F, sep = "\t")
#write.table(y.only.y.response.auto, "monocyte_autosomal_Ypresent_Yresponse.txt", quote = F, sep = "\t")

####Y-absent karyotypes model (X, XX, XXX)
y.absent.response <- GenomeWideResponse(mono.y.absent.dds, "X_count", expressedGenes)
y.absent.response.auto <- AutosomalResponse(y.absent.response)

#write.table(y.absent.response.auto, "monocyte_autosomal_Yabsent_Xresponse.txt", quote = F, sep = "\t")

####1 Y karyotypes model
y1_responsive <- GenomeWideResponse(mono.1y.dds, "X_count", expressedGenes)
y1_responsive.auto <- AutosomalResponse(y1_responsive)

#write.table(y1_responsive.auto, "monocyte_autosomal_1Y_Xresponse.txt", quote = F, sep = "\t")

####1 X, 1>=Y karyotypes model
x1_responsive <- GenomeWideResponse(mono.1x.dds, "Y_count", expressedGenes)
x1_responsive.auto <- AutosomalResponse(x1_responsive)

#write.table(x1_responsive.auto, "monocyte_autosomal_1X_Yresponse.txt", quote = F, sep = "\t")
```

###Adjust read counts
```{r}
mono.counts <- mono.txi$counts
adj.mono.cts <- batch_correct(mono.counts, colData)
adj.mono.cts <- adj.mono.cts$adj_cts
adj.mono.cts <- adj.mono.cts[rownames(adj.mono.cts) %in% expressedGenes,]

adj.mono.cts.meta <- merge(colData, t(adj.mono.cts), by = 0)
rownames(adj.mono.cts.meta) <- adj.mono.cts.meta$Row.names
adj.mono.cts.meta <- adj.mono.cts.meta[,-1]
adj.mono.cts.meta <- t(adj.mono.cts.meta)

norm.mono.counts <- counts(mono.dds, normalized = TRUE)
```

###Individual gene plots
```{r}
source("/sca-immune/Rfunctions/Function_ggGenePlot_regLine_xcount_auto.R")
source("/sca-immune/Rfunctions/Function_ggGenePlot_regLine_ycount_auto.R")
myPath <-  #set path to directory with deseq2 results
myOrange <- "#d95f02"
myOrange_light <- "#d95f0220"
myPurple <- "#7570B3"
myPurple_light <- "#7570B320"

setwd("") #set path to directory with deseq2 results
norm.cts <- counts(mono.dds, normalized=TRUE)
norm.cts <- norm.cts[rownames(norm.cts) %in% x.response$Gene,]
norm.cts.meta <- merge(colData, t(norm.cts), by = 0)

library(ggplot2)
library(doBy)
   
norm.cts.meta.df <- data.frame(norm.cts.meta)

ggGenePlot_regLine_xcount_auto("MVB12B", norm.cts.meta.df, normBy = 1000, prefix = "CD4", myWidth=1.25, myHeight=1.25)
ggGenePlot_regLine_ycount_auto("HCG11", norm.cts.meta.df, normBy = 1000, prefix = "CD4", myWidth=1.25, myHeight=1.25)

```

###Volcano plots
```{r}
myOrange <- "#d95f02"
myOrange_light <- "#d95f0220"
myPurple <- "#7570B3"
myPurple_light <- "#7570B320"

#Plot X-response -- volcano
#choose genes to highlight on volcano plot
examples <- c("MVB12B", "HCG11")

#set up significant genes, examples genes, and non-significant genes to be different colors
x.response.auto$significant <- "Not significant"
x.response.auto$significant[x.response.auto$padj < 0.05] <- "Significant response"
x.response.auto$significant[x.response.auto$Gene %in% examples] <- "Example gene"

#set up example genes to be labeled
x.response.auto$delabel <- NA
x.response.auto$delabel[x.response.auto$significant == "Example gene"] <- x.response.auto$Gene[x.response.auto$significant == "Example gene"]




library(ggrepel)
p <- ggplot(data = x.response.auto, aes(x = log2FoldChange, y = -log10(padj), col=significant, label = delabel)) +
  geom_point() +
  scale_color_manual(values=c("#D55E00", "#40404040", "#E69F0080")) +
  geom_text_repel(fontface = "italic") +
  labs(title = "Monocyte: autosomal response to Chr X", x="log2FC per Chr X") +
  geom_vline(xintercept=0, col="black") +
  geom_hline(yintercept=0, col="black")+
  ylim(NA,14) +
  theme_minimal(base_size = 7)+
  theme(legend.position="none")
p
ggsave("Figures/monocyte_autosomal_xresponse_volcano.pdf", p, width = 2, height = 2.75)


#Plot Y-response -- volcano
#set up significant genes, examples genes, and non-significant genes to be different colors
y.response.auto$significant <- "Not significant"
y.response.auto$significant[y.response.auto$padj < 0.05] <- "Significant response"
y.response.auto$significant[y.response.auto$Gene %in% examples] <- "Example gene"

#set up example genes to be labeled
y.response.auto$delabel <- NA
y.response.auto$delabel[y.response.auto$significant == "Example gene"] <- y.response.auto$Gene[y.response.auto$significant == "Example gene"]


p <- ggplot(data = y.response.auto, aes(x = log2FoldChange, y = -log10(padj), col=significant, label = delabel)) +
  geom_point() +
  scale_color_manual(values=c(myPurple, "#40404040", "#7570B360")) +
  geom_text_repel(fontface = "italic") +
  labs(title = "Monocyte: autosomal response to Chr Y", x="log2FC per Chr Y") +
  geom_vline(xintercept=0, col="black") +
  geom_hline(yintercept=0, col="black")+
  ylim(NA,14) +
  theme_minimal(base_size = 7)+
  theme(legend.position="none")
p
ggsave("Figures/monocyte_autosomal_yresponse_volcano.pdf", p, width = 2, height = 2.75)


```

###Compare X vs. Y response models
```{r}
overlap <- intersect(x.response.sig.auto$Gene,y.response.sig.auto$Gene)
xy_response_all_genes <- merge(x.response.auto, y.response.auto, by = "Gene")
write.table(xy_response_all_genes, "mono_x_and_y_responses.txt", quote = F, sep = "\t")
xy_response_all_genes <- read.csv("mono_x_and_y_responses.txt", sep ="\t")


#54 autosomal X-responsive genes
#46 autosomal Y-responsive genes
#11 autosomal X and Y-responsive overlapping genes
#12140 protein-coding and lincRNA expressed genes
phyper(11-1,54,12140-54,46,lower.tail= FALSE)

x_or_y <- xy_response_all_genes[xy_response_all_genes$padj.x < 0.05 | xy_response_all_genes$padj.y < 0.05,]

x_and_y <- xy_response_all_genes[xy_response_all_genes$padj.x < 0.05 & xy_response_all_genes$padj.y < 0.05,]

source("/sca-immune/Rfunctions/Function_makeSymmetric_X_Y.R")
source("/sca-immune/code/Rfunctions/Function_getMaxMin.R")
source("/sca-immune/Rfunctions/Function_paper2_corrPlot_deming_20230419.R")

#Deming regression with confidence intervals
##Monocyte X response vs. monocyte Y response
#all expressed genes
myData <- xy_response_all_genes[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = ".mono_Xresponse_v_Yresponse_autosomal_genes_deming_w_err.pdf", myXlab = "log2FC per Chr X", myYlab = "log2FC per Chr Y", myPlotTitle = "Monocyte: all autosomal genes", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)


#significant X- or Y-responsive genes
x_or_y$model <- "#6e02d980"
x_or_y$model[x_or_y$padj.x <0.05] <- "#D95F0280"
x_or_y$model[x_or_y$padj.x <0.05 & x_or_y$padj.y < 0.05] <- "#40404040"
myData <- x_or_y[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y", "model")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj", "model")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/mono_chrX_v_chrY_sig_err.pdf", myXlab = "log2FC per Chr X", myYlab = "log2FC per Chr Y", myPlotTitle = "Monocyte: Chr X or Chr Y responsive", myWidth=2.25, myHeight=2.25, pointsCol = myData$model)


x_or_y$significance <- "X-response only (42 genes)"
x_or_y$significance[x_or_y$padj.y < 0.05] <- "Y-response only (31 genes)"
x_or_y$significance[x_or_y$padj.x < 0.05 & x_or_y$padj.y < 0.05] <- "Both (10 genes)"

#some stats
x.only.sig <- x_or_y[x_or_y$significance == "X-response only (42 genes)",]
cor.test(x.only.sig$log2FoldChange.x, x.only.sig$log2FoldChange.y)$p.value
cor.test(xy_response_all_genes$log2FoldChange.x, xy_response_all_genes$log2FoldChange.y)$p.value


```

#X vs. Y responsive Venn diagram
```{r}
x_y.auto <- list(x.response.sig.auto$Gene, y.response.sig.auto$Gene)
names(x_y.auto) <- c("Chr X responsive", "Chr Y responsive")
pdf(file = "mono.x_vs_y_autosomal_venn.pdf", width = 4, height = 3)
plot(euler(x_y.auto), quantities = TRUE)
dev.off()

```

#Compare all samples model ("full model") vs. Y-present model
```{r}
#Y-response, comparing model using all samples versus model with just Y karyotypes
all.y.present.xresponse <- merge(x.response.auto, y.only.x.response.auto, by = "Gene")
all.y.present.yresponse <- merge(y.response.auto, y.only.y.response.auto, by = "Gene")

cor.test(all.y.present.xresponse$log2FoldChange.x, all.y.present.xresponse$log2FoldChange.y, method = "pearson")$p.value
cor.test(all.y.present.yresponse$log2FoldChange.x, all.y.present.yresponse$log2FoldChange.y, method = "pearson")$p.value

all.y.present.xresponseSig <- merge(x.response.sig.auto, y.only.x.response.auto, by = "Gene")
all.y.present.yresponseSig <- merge(y.response.sig.auto, y.only.y.response.auto, by = "Gene")

cor.test(all.y.present.xresponseSig$log2FoldChange.x, all.y.present.xresponseSig$log2FoldChange.y, method = "pearson")
cor.test(all.y.present.yresponseSig$log2FoldChange.x, all.y.present.yresponseSig$log2FoldChange.y, method = "pearson")


##Deming confidence intervals
source("/sca-immune/Rfunctions/Function_makeSymmetric_X_Y.R")
source("/sca-immune/Rfunctions/Function_getMaxMin.R")
source("/sca-immune/Rfunctions/Function_paper2_corrPlot_deming_20230419.R")

myData <- all.y.present.xresponseSig[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/monocyte.full_vs_Ypresent_model_XresponseAutoSigGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr X", myYlab = ">1 Y model, log2FC per Chr X", myPlotTitle = "Monocyte: Full model vs. >1 Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)


myData <- all.y.present.yresponseSig[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/monocyte.full_vs_Ypresent_model_YresponseAutoSigGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr Y", myYlab = ">1 Y model, log2FC per Chr Y", myPlotTitle = "Monocyte: Full model vs. >1 Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)

###all expressed genes
myData <- all.y.present.xresponse[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/monocyte.full_vs_Ypresent_model_XresponseAutoGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr X", myYlab = ">1 Y model, log2FC per Chr X", myPlotTitle = "Monocyte: Full model vs. >1 Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)


myData <- all.y.present.yresponse_to_plot[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/monocyte.full_vs_Ypresent_model_YresponseAutoGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr Y", myYlab = ">1 Y model, log2FC per Chr Y", myPlotTitle = "Monocyte: Full model vs. >1 Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)


```

####Compare all samples model ("full model") vs. Y-absent ("no Y") model
```{r}
all.y.absent.xresponse <- merge(x.response.auto, y.absent.response.auto, by = "Gene")
cor.test(all.y.absent.xresponse$log2FoldChange.x, all.y.absent.xresponse$log2FoldChange.y, method = "pearson")
cor.test(all.y.absent.xresponse$log2FoldChange.x, all.y.absent.xresponse$log2FoldChange.y, method = "pearson")$p.value

all.y.absent.xresponseSig <- merge(x.response.sig.auto, y.absent.response.auto, by = "Gene")
cor.test(all.y.absent.xresponseSig$log2FoldChange.x, all.y.absent.xresponseSig$log2FoldChange.y, method = "pearson")$p.value


##Deming confidence intervals
source("/sca-immune/Rfunctions/Function_makeSymmetric_X_Y.R")
source("/sca-immune/Rfunctions/Function_getMaxMin.R")
source("/sca-immune/Rfunctions/Function_paper2_corrPlot_deming_20230419.R")

###significantly Chr X responsive genes
myData <- all.noY.xresponseSig[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/mono.full_vs_noY_model_XresponseSigAutoGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr X", myYlab = "no Y model, log2FC per Chr X", myPlotTitle = "Monocyte: Full model vs. no Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)

###all autosomal expressed genes
myData <- all.noY.xresponse[,c("Gene","log2FoldChange.x", "lfcSE.x", "padj.x", "log2FoldChange.y", "lfcSE.y", "padj.y")]
colnames(myData) <- c("Gene","x","x_err","x_padj", "y", "y_err", "y_padj")
paper2_corrPlot_deming_err(myData, myTitle = "Figures/mono.full_vs_noY_model_XresponseAutoGenes_deming_err.pdf", myXlab = "Full model, log2FC per Chr X", myYlab = "no Y model, log2FC per Chr X", myPlotTitle = "Monocyte: Full model vs. no Y model", myWidth=2.25, myHeight=2.25, demingBehind = TRUE)


```

